Theories of deliberated choice

### 2023-11-23 simple.tex
Progress towards indicativeness

Given an observation $\theta$, let $\theta^{-1}(P) \subseteq I × \allhists$ denote the pairs $(i, \hist)$ in $\theta$.
Let $\incr(\allhists) = \set{\emptyset} \cup \cup_{\emptyset \neq N \subseteq \N} \set{A(N) \suchthat A \in (\allhists)^N \land [\card{A_i} = i] \land [i < j \implies A_i \subsetneq A_j]} \subseteq \powerset{\allhists}$ denote the possibly empty embedded sequences of paths.
An element $\theta^{-1}(P)$ can also be seen as a (finite) function associating to each $i \in I$ an element of $\incr(\allhists)$.
Therefore, we let $\finite(\incr(\allhists)^I) = \set{\theta^{-1}(P) \suchthat \theta \in \Theta}$ denote the set of possible observations restricted to $I × \allhists$, the elements of which being called \emph{a priori observation}, where $\finite(\incr(\allhists)^I)$ denote the functions $f: I \mapsto \incr(\allhists)$ with finite support, thus $\card{f^{-1}(\incr(\allhists) \setminus \emptyset)} < \N$.
Note that $\allpriobss$ and $\incr(\allhists)^I$ are downwards closed, meaning that $\forall O \in \allpriobss: O' \subseteq O \implies O' \in \allpriobss$ and $\forall O \in \incr(\allhists)^I: O' \subseteq O \implies O' \in \incr(\allhists)^I$.
Let also $\max\incr(\allhists) = \set{A \in \incr(\allhists) \suchthat \nexists A' \in \incr(\allhists) \suchthat A \subset A'} = \set{A \in \incr(\allhists) \cap (\allhists)^\N}$ denote the maximal elements of $\incr(\allhists)$, in the sense of containing sequences that are not a subsequence of another sequence; equivalently, containing those sequences whose sequence of cardinals of their elements equals $\N$.

@@

\begin{remark}
  \Cref{ex:threeFs} provides an example where, similarly, the claim may fail to hold but still be not rejectable; but where rejection depends on ${\allleadsto}$.
\end{remark}

\Cref{def:falsw} fills the gap to obtain that $C$ holds iff it is not rejectable.
The property demands that whatever the attitude of individuals, if the attitude is not according to the claim, then it is possible to observe that it is not.
\begin{definition}
  \label{def:falsw}
  A claim $C$ is \emph{falsifiable or trivial} iff
  $\forall {\allleadstog} \notin C: \exists θ \in \Theta \cap \powerset{\allleadstog} \suchthat \theta \notin C^\wsub$.
\end{definition}

@@

\begin{definition}
  A claim is \emph{strongly unfalsifiable} iff
  $\Theta \wsub C^\wsub$.
\end{definition}

\begin{example}[Big disjunct, strongly unfalsifiable]
  \label{ex:bigor}
  Consider $\ar \in \allargs$.
  The claim $[\forall i: (\ar) \ileadstog \phi] \lor [\forall i: (\ar, \ar) \ileadstog \phi] \lor …$,
  thus, $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$,
  is strongly unfalsifiable.

  Given any $θ \in \Theta$, define $l = \max_{\hist \suchthat \exists i \in I, p \in P \suchthat (i, \hist, p) \in θ} \card{\hist} + 1$ if $\exists p \in P \suchthat (i, \hist, p) \in θ$, and $l = 1$ otherwise. Pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have
  $\theta \in \powerset{\allleadstog}$ and ${\allleadstog} \in C$, thus, $\theta \in C^\wsub$.
\end{example}
\begin{example}[Disjunct with incompatible paths, not falsifiable]
  \label{ex:disj}
  Given $\ar \neq \ar' \in \allargs$, the claim appearing in \cref{ex:disjR}, namely,
  $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$, is not falsifiable.
  Consider any ${\allleadstog} \notin C$, thus satisfying, for some $i$, $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
  Consider any $θ \subseteq {\allleadstog}$ and let us prove that $\theta \in C^\wsub$.
  By definition of a general observation, $θ$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $θ$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have 
  $\theta \in \powerset{\allleadstog'}$ and ${\allleadstog'} \in C$, thus, $\theta \in C^\wsub$.
\end{example}
\begin{example}[Partly falsifiable is still unfalsifiable]
  \label{ex:partlyU}
  Given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
  $\hist \ileadstog[j] \phi \land \forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
  Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
  \Cref{ex:disj} concludes.
\end{example}
\begin{example}
  \label{ex:allParticular}
  Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
  Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $θ = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
  \label{ex:existsParticular}
  Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
  Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $θ \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $\theta \in C^\wsub$.
\end{example}

\begin{remark}
  The claim “For each glass, it will not break on a monday or not break on a tuesday, whatever happens to it.” is not falsifiable. But one can make it stronger and falsify it, by giving a general theory of how glasses break, proving that it holds, and that it is not compatible with the claim. I should think about this.
\end{remark}
Let $S_θ = \set{{\allleadstog} \in \allleadstos \suchthat θ \subseteq {\allleadstog}}$ denote the possibilities for $\allleadsto$ compatible with the observations.
Given $O$ and ${\allleadstog}$, it follows that $S_{\restr{\allleadstog}{O_O}}$ denotes the “leads to” relations that complete $O_O$ using $\allleadstog$.
\begin{remark}
  To be falsifiable, I need each time I exclude some relation $R$ to also exclude a subset $Cbar$ such that some $O \subseteq R$ satisfies $S_O$ is entirely in $Cbar$.
\end{remark}
\begin{theorem}
  A claim $C$ is falsifiable iff its complement $\Cbar = \allleadstos \setminus C$ satisfies:
  $\forall {\allleadstog} \in \Cbar, \exists \theta \subseteq {\allleadstog} \suchthat S_θ \subseteq \Cbar$, or equivalently:
  $\forall {\allleadstog} \in \Cbar, \exists O \in (\allhists)^I \suchthat S_{\restr{\allleadstog}{O_O}} \subseteq \Cbar$.
\end{theorem}
\begin{proof}
  The first phrasing follows from the definition of falsifiability.
  Also, the second phrasing implies the first one: pick $\theta = \restr{\allleadstog}{O_O}$.

\begin{theorem}
  \label{thm:caractFals}
  A claim $C$ is falsifiable iff $\forall {\allleadstog} \in \allleadstos$:
  \[\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq \implies {\allleadstog} \in C.\]
\end{theorem}
\begin{proof}
  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: \theta \in C^\wsub$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: θ \in C^\subseteq$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.

  To conclude, note that $\forall {\allleadstog} \in \allleadstos$: $\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} = \set{θ \subseteq {\allleadstog}}$.
\end{proof}
\begin{corollary}
  A claim $C$ is falsifiable iff $\forall f \in C^{\allpriobss} \suchthat \forall O, O' \in \allpriobss: \restr{f(O)}{O \cap O'} = \restr{f(O')}{O \cap O'}$:
  \[\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C.\]
\end{corollary}
\begin{proof}
  Consider a falsifiable claim $C$, consider such an $f$, and let us prove that $\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C$.
  Define ${\allleadstog} \subseteq (I × \allhists) × P$ as $\cup_{O \in \allpriobss} \restr{f(O)}{O}$. Its domain equals $I × \allhists$, and it is right-unique thanks to the condition on $f$.
  Thus, \cref{thm:caractFals} applies and ${\allleadstog} \in C$.

We thus reach a procedure to determine whether a claim is falsifiable.
Given $C$, compute $C^\subseteq$. Find the set $R$ of relations $\allleadstog$ satisfying $\forall O: \restr{\allleadstog}{O} \in C^\subseteq$. The claim $C$ is falsifiable iff $R \subseteq C$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq} \subseteq C$.
\end{corollary}
A relation $\allleadstog$ is observably in $C$ iff $\forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq$, equivalently, iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)$.
All relations in $C$ are observably in $C$. The question of falsifiability asks whether the converse holds.
In other words, $C$ is falsifiable iff it includes the relations observably in $C$.
Define that a relation ${\allleadstog} \in {\allleadstos}$ is \emph{obs-compatible} with a set $S$ iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S$.
Then, the relations observably in $C$ are exactly those obs-compatible with $C^\subseteq \cap \Theta$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)} = C$.
\end{corollary}
\begin{remark}
  The question of falsifiability is whether the set of relations $\allleadstog$ “compatible” with $C^\subseteq \cap \Theta$ equals $C$.
  Compatibility needs to consider the observables, not a subset of it.
  Define $\Theta_1 \subset \Theta$ as the observations that observe only one pair $(i, \hist)$.
  Define $R_1 = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta_1: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)}$.
  Note that $R \subseteq R_1$.
  We know that $C$ is falsifiable iff $R \subseteq C$. Thus, $C$ is not falsifiable implies $C \subsetneq R \subseteq R_1$ thus $C \subsetneq R_1$. But we do not have that $C \subsetneq R_1$ implies $C$ not falsifiable: could have $C \subsetneq R_1$ but still $R = C$.
\end{remark}
\begin{example}[Come back]
  Considering again \cref{ex:bigor}, by \cref{th:veryunfals}, we see that $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$ is not falsifiable.

  Similarly, considering again \cref{ex:disj}, given $\ar \neq \ar' \in \allargs$,
  $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable, by \cref{th:veryunfals}.

  Turning to \cref{ex:partlyU},
  given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
  $C = \set{{\allleadstog} \suchthat \hist \ileadstog[j] \phi \land \forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]}$ is not falsifiable but \cref{th:veryunfals} is not applicable.
  Observe that
  $C^\subseteq = \set{R \in I × \allhists × P \suchthat [(j, \hist) \in R^{-1}(P) \implies (j, \hist, \phi) \in R] \land \forall i: [(i, (\ar)) \in R^{-1}(P) \land (i, (\ar')) \in R^{-1}(P)] \implies [(i, (\ar), \phi) \in R \land (i, (\ar'), \phi) \in R]}$ and
  $C^\subseteq \cap \Theta = \set{\theta \in \Theta \suchthat (j, \hist) \in \theta^{-1}(P) \implies (j, \hist, \phi) \in \theta}$.
  % Pick any $\allleadstog$ satisfying $\hist \ileadstog[j] \phi$ and $\exists i \suchthat (\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
  % The definition of non-falsifiability applies: pick any $\theta \subseteq {\allleadstog}$, we have $\theta \in C^\subseteq$ as $(j, \hist, \phi)$ is compatible with $\theta$ and the other condition excludes nothing in $\Theta$.
  % We can apply Thm 4 similarly: given any $O$, we have $\restr{\allleadstog}{O} \in C^\subseteq$.
  % We should be able to apply a corollary similarly.
  The relations obs-compatible with $C^\subseteq \cap \Theta$ are those satisfying $\hist \ileadstog[j] \phi$.
  We conclude that $C$ is not falsifiable as $C$ is smaller than that set.

  In \cref{ex:allParticular} we considered a given $\ar$ and $\phi$ and defined the falsifiable claim $C = \set{{\allleadstog} \suchthat \forall i: \ar \ileadstostg \phi}$.
  Observe that $C^\subseteq \cap \Theta = \set{\theta \in \Theta \suchthat \forall \hist \suchthat \ar \in \histend: \theta(i, \hist) = \emptyset \lor \theta(i, \hist) = \phi}$.
  The relations obs-compatible with $C^\subseteq \cap \Theta$ are exactly those in $C$.

\begin{remark}
  \label{rq:closedcomb}
  Being closed-under-combinations (where a combination forms a new relation by union of parts of relations in $C$) is sufficient to be falsifiable, but not necessary, as seen with \cref{ex:consti2}.
\end{remark}

\begin{remark}
  Intuition (to be explored, TODO) is that when $C^\wsub \cap \Theta$ has the covering property, then the obs-compatible relations are easily deduced from the set (something like, just consider compatibility point per point, one singleton observation at a time). Covering is what is used above, something like $\exists i, \hist, \histp$ such that $\forall \hist^{\prime\prime}: [\forall j \neq i: \set{(\hist, i, .), (\histp, j, .)} \in C^\wsub \cap \Theta] \land [\exists j \neq i \suchthat \set{(\hist', i, .), (\hist^{\prime\prime}, j, .)} \in C^\wsub \cap \Theta]$.
\end{remark}

@@

Given an exposure protocol $\allleadsto$, deliberated choices are represented by an element $\pi({\allleadsto}) \subseteq I × P$ which is right unique (meaning that $\forall i \in I: \card{\pi({\allleadsto})(i)} ≤ 1$).
Equivalently, let $(P \cup \set{\emptyset})^I$ denote the right unique relations over $I × P$; and then the deliberated choices are an element $\pi({\allleadsto}) \in (P \cup \set{\emptyset})^I$.
We say that $C$ is \emph{informative} iff $\exists {\allleadstog} \in (P \cup \set{\emptyset})^I \suchthat C \implies ¬{\allleadstog}$, meaning that $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.

@@

>>>>

Given an observation $\theta$, let $\theta^{-1}(P) \subseteq I × \allhists$ denote the pairs $(i, \hist)$ in $\theta$.
Let $\incr(\allhists) = \set{\emptyset} \cup \cup_{\emptyset \neq N \subseteq \N} \set{A(N) \suchthat A \in (\allhists)^N \land [\card{A_i} = i] \land [i < j \implies A_i \subsetneq A_j]} \subseteq \powerset{\allhists}$ denote the possibly empty embedded sequences of paths.
\commentOC{TODO define $\incr(\allhists) = \set{H \wsub \allhists \suchthat \card(H) ≤ \card(H') \implies H \wsub H'}$.}
An element $\theta^{-1}(P)$ can also be seen as a (finite) function associating to each $i \in I$ an element of $\incr(\allhists)$.
Therefore, we let $\finite(\incr(\allhists)^I) = \set{\theta^{-1}(P) \suchthat \theta \in \Theta}$ denote the set of possible observations restricted to $I × \allhists$, the elements of which being called \emph{a priori observation}, where $\finite(\incr(\allhists)^I)$ denote the functions $f: I \mapsto \incr(\allhists)$ with finite support, thus $\card{f^{-1}(\incr(\allhists) \setminus \emptyset)} < \N$.
Note that $\allpriobss$ and $\incr(\allhists)^I$ are downwards closed, meaning that $\forall O \in \allpriobss: O' \subseteq O \implies O' \in \allpriobss$ and $\forall O \in \incr(\allhists)^I: O' \subseteq O \implies O' \in \incr(\allhists)^I$.
Let also $\max\incr(\allhists) = \set{A \in \incr(\allhists) \suchthat \nexists A' \in \incr(\allhists) \suchthat A \subset A'} = \set{A \in \incr(\allhists) \cap (\allhists)^\N}$ denote the maximal elements of $\incr(\allhists)$, in the sense of containing sequences that are not a subsequence of another sequence; equivalently, containing those sequences whose sequence of cardinals of their elements equals $\N$.

@@

\begin{remark}
  \Cref{ex:bigor,ex:threeFs} present situations where, similarly, the claim may fail to hold but still be not rejectable.
  In \cref{ex:bigor}, non rejection is due to finiteness of observations.
  In \cref{ex:threeFs}, rejection depends on ${\allleadsto}$.
\end{remark}

\Cref{def:falsw} fills the gap to obtain that $C$ holds iff it is not rejectable.
The property demands that whatever the attitude of individuals, if the attitude is not compatible with the claim, then it is possible to observe that it is not.
\begin{definition}
  \label{def:falsw}
  A claim $C$ is \emph{falsifiable or trivial} iff
  $\forall {\allleadstog} \notin C: \Theta \cap \powerset{\allleadstog} \nsubseteq C^\wsub$.
\end{definition}

@@

\begin{definition}
  A claim is \emph{strongly unfalsifiable} iff
  $\Theta \wsub C^\wsub$.
\end{definition}
\Cref{ex:disjR,ex:bigor} illustrate, as well as \cref{ex:existsParticular} when $I$ is infinite.
\begin{definition}
  A claim is \emph{strictly unfalsifiable} iff
  it is unfalsifiable and not strongly unfalsifiable.
\end{definition}

Making a strongly unfalsifiable claim more precise ($C = (C_1 \cap C_2) \sub C_1$ in \cref{th:strongUToStrictU}) can make it strictly unfalsifiable.
Also, a strictly unfalsifiable claim can be made falsifiable by picking a single relation from the claim.

But, it does not hold that more precise claims are always more falsifiable:
making a falsifiable claim more precise ($C = (C_1 \cap C_2) \sub C_2$ in \cref{ex:partlyU}) can make it strictly unfalsifiable.
(In fact, \cref{ex:partlyU} shows that the relation is not one-way: going from a strongly unfalsifiable ($C_1$), or from a falsifiable ($C_2$), claim, to a more precise one ($C = C_1 \cap C_2$) yields a strictly unfalsifiable claim.)

\begin{example}
  \label{ex:allParticular}
  Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
  Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $θ = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}

The following procedure permits to determine whether a claim is falsifiable.
Given a set of observations $S \wsub \Theta$, define the relations that are observation-compatible with $S$ as $E(S) = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: {\allleadstog} \cap \theta \in S}$.
Given a claim $C$, define the set of its observation-compatible relations as $E(C) = E(C^\wsub \cap \Theta)$ (reusing the letter $E$ by abuse of notation).
Note that this set extends $C$, in the sense that $C \wsub E(C)$.
The claim $C$ is falsifiable iff it is closed under extension to observation-compatible relations, that is, iff $E(C) \subseteq C$.
\begin{theorem}
  \label{thm:caractFals}
  A claim $C$ is falsifiable or trivial iff $E(C) \wsub C$, or equivalently, iff 
  $\forall {\allleadstog} \in \allleadstos$:
  \[\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq \implies {\allleadstog} \in C.\]
\end{theorem}
\begin{proof}
  Note that
  $E(C) = E(C^\wsub \cap \Theta) = \set{{\allleadstog} \in \allleadstos \suchthat \forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\wsub}$.

  The condition $E(C) \wsub C$ is equivalent to
  $\forall {\allleadstog} \in \allleadstos: ({\allleadstog} \in E(C) \implies {\allleadstog} \in C)$.
  Note that ${\allleadstog} \in E(C) \Leftrightarrow \set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq$.

  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \Theta \cap \powerset{\allleadstog} \wsub C^\wsub$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.
  And, $\forall {\allleadstog} \in \allleadstos$: $\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} = \set{θ \subseteq {\allleadstog}}$.
\end{proof}

Let $S_θ = \set{{\allleadstog} \in \allleadstos \suchthat θ \subseteq {\allleadstog}}$ denote the possibilities for $\allleadsto$ compatible with the observations.
Given $O$ and ${\allleadstog}$, it follows that $S_{\restr{\allleadstog}{O_O}}$ denotes the “leads to” relations that complete $O_O$ using $\allleadstog$.
Define that a relation ${\allleadstog} \in {\allleadstos}$ is \emph{obs-compatible} with a set $S$ iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S$.

\begin{remark}
  \label{rq:closedcomb}
  Define the closure under combinations of $C$ as the relations that can be formed by picking parts of relations in $C$, formally, $(\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos$.
  We have $E(C) \wsub (\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos$.

  Being closed under combinations is sufficient to be falsifiable, but not necessary, as seen with \cref{ex:consti2} (in other words, for some claims, $(\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos \wsub E(C)$ does not hold).
\end{remark}

@@

Given an attitude $\allleadsto$, deliberated choices are represented by an element $\pi({\allleadsto}) \subseteq I × P$ (thus with $\pi({\allleadsto})(i) = P_i$) which is right unique (meaning that $\forall i \in I: \card{\pi({\allleadsto})(i)} ≤ 1$).
Equivalently, let $(P \cup \set{\emptyset})^I$ denote the right unique relations over $I × P$; and then the deliberated choices are an element $\pi({\allleadsto}) \in (P \cup \set{\emptyset})^I$.

A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
% A claim $C$ is \emph{indicative} iff $\exists H \in \finite(\incr(\allhists)), i \in I \suchthat \pi(\set{{\allleadstog} \suchthat H \wsub \ileadstoinvg(P)} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub \suchthat \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
\begin{remark}
  Informativeness permits mere counting: the claim that exactly two persons deliberately believe $\phi$ is informative.
  Informativeness may be a desirable property (rather than informativeness beyond counting) as one may want to know whether a given proposition is believed by a majority of individuals, or whether a given proposition is believed by nobody.
  Note also that strong informativeness beyond counting implies informativeness beyond counting and indicativeness.
  Note also that informativeness beyond counting implies informativeness.
  \Cref{th:infoBeyonCKO} (and extensions, \cref{th:infoBeyonCKOext,th:strinfoBeyonCKO}) shows that informativeness beyond counting (and variants) is too demanding if one wants to investigate situations where individuals have differing deliberate preferences.
\end{remark}

@@

\begin{theorem}
  For all exposure protocols, there exists a claim that holds and is strongly informative beyond counting and falsifiable iff $\exists i \suchthat P_i \neq \emptyset$.
\end{theorem}
\begin{proof}
  Define $C = \set{\allleadsto}$.
\end{proof}
\begin{theorem}
  For all exposure protocols, there exists a claim that holds and is informative beyond counting and falsifiable.
\end{theorem}
\begin{proof}
  Define $C = \set{\allleadsto}$.
\end{proof}
\begin{theorem}
  For all exposure protocols, there exists a claim that is anonymous, strongly informative beyond counting and falsifiable.
\end{theorem}
\begin{proof}
  Define the claim as $\forall i \in I, \hist \in \allhists: \hist \ileadstog \phi$.
\end{proof}

What about anonymous and holds?
\begin{theorem}
  For all exposure protocols, there exists a claim that holds and is anonymous.
\end{theorem}
\begin{proof}
  Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}$.
\end{proof}

\begin{theorem}
  Assume that $I = \set{i_1, i_2}$ and that $P_{i_1} = \phi$ and $P_{i_2} = ¬\phi$.

  Then, there exists a claim that holds and is anonymous and indicative.
\end{theorem}
\begin{proof}
  From $P_{i_1} = \phi$ we know that some $a_1$ is a decisive argument in favor of $\phi$ for $i_1$.
  From $P_{i_2} = ¬\phi$ we know that for some $\hist \in \allhists \suchthat a_1 \in \histend$, $\hist \nileadsto[i_2] \phi$ whereas $\hist \ileadsto[i_1] \phi$.
  Define $O = \set{(\hist, i_1)}$.
  Define 
  ${\allleadstog^1} = \set{(i_1, \ileadsto[i_1]), (i_2, \ileadsto[i_2])}$ and
  ${\allleadstog^2} = \set{(i_1, \ileadsto[i_2]), (i_2, \ileadsto[i_1])}$.
  Define the claim as $C = \set{{\allleadstog^1}, {\allleadstog^2}}$.
\end{proof}

\begin{theorem}
  Assume that $I = \set{i_1, i_2, i_3}$ and that $P_{i_1} = \phi$ and $P_{i_2} = P_{i_3} = ¬\phi$.
  Assume that all paths lead either to $\phi$ or to $¬\phi$.
  Define $H_2$ as the set of paths that lead $i_2$ to $¬\phi$ and similarly $H_3$ for $i_3$.
  In other words,
  $\forall \hist \in \allhists \setminus H_2: \hist \ileadsto[i_2] \phi$,
  $\forall \hist \in \allhists \setminus H_3: \hist \ileadsto[i_3] \phi$,
  $\forall \hist \in H_2: \hist \ileadsto[i_2] ¬\phi$,
  $\forall \hist \in H_3: \hist \ileadsto[i_3] ¬\phi$.
  Also, define $H_{12}$ as the set of paths on which $i_1$ and $i_2$ differ.
  Similarly for $H_{13}$.
  We want small such sets, so let us assume that $i_1$ goes often to $¬\phi$ and only on some $H_1$ including all paths that ends with $a_1$ does it go to $\phi$.
  We need to have suffix compatible observations from $H_{12}$ and $H_{13}$.
  How to make none such?
  Impossible: consider $(a_1, a_2)$ and $(a_1, a_2, a_1, a_3)$.

  Thus with $H_2$ including all paths that ends with some $a_2$ decisive for $i_2$.
  Thus, the path $(a_2, a_3)$ is in both $H_2$ and $H_3$.

  Then, there exists a claim that holds and is anonymous and indicative?

  But maybe $i_2$ always goes to $\phi$ except for the paths $H_2$, and $i_3$ always goes to $\phi$ except for the paths $H^3$ incompatible (not suffix or prefix) with $H_2$.
  Then we can’t observe both that $i$ is not $i_2$ and not $i_3$: whatever $O$, if we fall upon $i_1$, this will leave $i_1$ and some other $i$ as possibilities.
\end{theorem}
\begin{theorem}
  Assume that $I = \set{i_1, i_2, i_3}$ and that $P_{i_1} = \phi$ and $P_{i_2} = P_{i_3} = ¬\phi$.

  Then, there exists a claim that holds and is anonymous and indicative.
\end{theorem}
\begin{proof}
  Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}$.
  In other words, for $k, l \in \intvl{1, 3}$, define
  ${\allleadstog^{k, l}} = \set{(i_k, \ileadsto[i_1]), (i_l, \ileadsto[i_2]), (i_{\intvl{1, 3} \setminus \set{k, l}}, \ileadsto[i_3])}$.
  Define the claim as $C = \set{{\allleadstog^{1, 2}}, {\allleadstog^{1, 3}}, {\allleadstog^{2, 1}}, {\allleadstog^{2, 3}}, {\allleadstog^{3, 1}}, {\allleadstog^{3, 2}}}$.

  Define $a_1$ as an argument decisive in favor of $\phi$ for $i_1$ and $a_2, a_3$ as arguments decisive in favor of $¬\phi$ for $i_2, i_3$.

  Define $\hist_{12} = (a_1, a_2)$ and $\hist_{13} = (\hist_{12}, a_1, a_3)$.
  We know that 
  $\hist_{12} \ileadsto[i_1] \phi$, 
  $\hist_{12} \ileadsto[i_2] ¬\phi$, 
  $\hist_{13} \ileadsto[i_1] \phi$, 
  $\hist_{13} \ileadsto[i_3] ¬\phi$.
  Define $\theta = \set{(\hist_{12}, i_1, \phi), (\hist_{13}, i_1, \phi)}$.
  Finally, the attitudes in $C$ compatible with $\theta$ are only $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C = \set{{\allleadstog^{1, 2}}, {\allleadstog^{1, 3}}}$, and $\pi(\set{{\allleadstog^{1, 2}}, {\allleadstog^{1, 3}}})(i_1) = \phi$.
\end{proof}

\begin{theorem}
  Assume that $P_{i_1} = \phi$ and $\forall i \in I \setminus \set{i_1}: P_i = ¬\phi$.

  Then, there exists a claim that holds and is anonymous and indicative.
\end{theorem}
\begin{proof}
  From $P_{i_1} = \phi$ we know that some $a_1$ is a decisive argument in favor of $\phi$ for $i_1$.
  For $i_1 \neq i \in I$, from $P_i = ¬\phi$ we know that for some $\hist^i \in \allhists \suchthat a_1 \in \histend[i]$, $\hist^i \nileadsto \phi$ whereas $\hist^i \ileadsto[i_1] \phi$.
  If $I$ is finite, define $O = \set{(\hist^i, i_1) \suchthat i_1 \neq i \in I}$.
  TODO this fails because $O$ might not be an a priori observation.
  Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto} \suchthat i \in I}$.
\end{proof}

\begin{theorem}
  \label{th:HAinformativeness}
  Given any $Q \wsub P \cup \set{\emptyset}$,
  the following two conditions on $\allleadsto$ are equivalent:
  there exists a claim that holds and is anonymous and satisfies 
  $\exists j \in I \suchthat \pi(C)(j) \ssub Q$;
  and 
  $\set{P_i, i \in I} \ssub Q$.
\end{theorem}
\begin{proof}
  Consider any claim $C$ that holds and is anonymous, pick any $j \in I$, and let us show that $\set{P_i, i \in I} \wsub \pi(C)(j)$.
  Pick any $p \in \set{P_i, i \in I}$. 
  Pick any $i \in I$ such that $\pi({\allleadsto})(i) = p$.
  Define $\sigma$ as the permutation exchanging $i$ and $j$.
  As $C$ holds, it contains $\allleadsto$, and by anonymity, it contains ${\allleadsto} \circ \sigma$. Observing that $\pi({\allleadsto} \circ \sigma)(j) = \pi({\allleadsto})(i) = p$, we obtain $p \in \pi(C)(j)$.

  It follows that for any claim $C$ that holds and is anonymous, if $\exists j \in I \suchthat \pi(C)(j) \ssub Q$ then $\set{P_i, i \in I} \wsub \pi(C)(j) \ssub Q$ thus $\set{P_i, i \in I} \ssub Q$.

  Turning to the other direction, if $\set{P_i, i \in I} \ssub Q$, then pick any $p \in Q \setminus \set{P_i, i \in I}$; the claim $\forall i \in I: P_i \neq p$ satisfies the desired properties.
\end{proof}

\begin{theorem}
  \label{th:infoBeyonCKO}
  There exists a claim that holds and is anonymous and informative beyond counting iff $\set{P_i, i \in I} \neq P \cup \set{\emptyset}$.
\end{theorem}
\begin{proof}
  From \cref{th:HAinformativeness} with $Q = P \cup \set{\emptyset}$, we have that 
  there exists a claim that holds and is anonymous and satisfies 
  $\exists j \in I \suchthat \pi(C)(j) \ssub P \cup \set{\emptyset}$ iff
  $\set{P_i, i \in I} \ssub P \cup \set{\emptyset}$.
\end{proof}
